Photo-D with LSST: stellar photometric distances out to the edge of the Galaxy

### 2024-03-07 paper/intro.tex
Paper  comments.
new sections 2.2.3 and 2.2.4, and comments from Lovro and Bob

As demonstrated with SDSS \citep{2008ApJ...673..864J, 2008ApJ...684..287I}, Pan-STARRS \citep{2014ApJ...783..114G, green_3d_2019} and most recently with
Gaia data \citep{bailer-jones_estimating_2021}, broad-band near-UV to near-IR stellar photometry is sufficient
to estimate distance, metallicity and interstellar dust extinction along the line of sight for stars in the Galaxy.
In analogy with photo-z, photometric redshifts of galaxies, hereafter we refer to these methods as photo-D. 
The photo-D method is conceptually quite simple: multi-dimensional color tracks (either empirical or model-based),
parametrized by luminosity, metallicity and extinction, are fit to observed colors and the best-fit produces estimates
of the three model parameters. The method relies on strong correlations between stellar colors and
stellar luminosity for dominant stellar populations such as main-sequence stars, red giants, white dwarfs, and even
for the majority of unresolved binary stars. These are the same correlations that are responsible for the abundant
structure seen in the Hertzsprung-Russell diagram. In addition to colors and luminosity, these correlations
also involve metallicity in case of main-sequence and red giant stars, and surface gravity in case of white dwarfs.
For the youngest main-sequence stars, stellar age may play a role, too. Measured stellar colors are also affected
by interstellar dust extinction along the line of sight towards the star. Consequently, sufficiently accurate measurements of apparent
brightness and sufficient number of UV to IR colors, such as those that Rubin Observatory's Legacy Survey of Space
and Time (LSST; \citealt{2019ApJ...873..111I}) will provide, can be used to accurately estimate these parameters,
and ultimately stellar distances.

>>>>

As demonstrated with SDSS \citep{2008ApJ...673..864J, 2008ApJ...684..287I}, Pan-STARRS \citep{2014ApJ...783..114G, green_3d_2019} and most recently with
Gaia data \citep{bailer-jones_estimating_2021}, broad-band near-UV to near-IR stellar photometry is sufficient
to estimate distance, metallicity and interstellar dust extinction along the line of sight for stars in the Galaxy.
In analogy with photo-z, photometric redshifts of galaxies, hereafter we refer to these methods as photo-D. 
The photo-D method is conceptually quite simple: multi-dimensional color tracks (either empirical or model-based),
parametrized by luminosity, metallicity and extinction, are fit to observed colors and the best-fit produces estimates
of the three model parameters. The method relies on strong correlations between stellar colors and
stellar luminosity for dominant stellar populations such as main-sequence stars, red giants, white dwarfs, and even
for the majority of unresolved binary stars. These are the same correlations that are responsible for the abundant
structure seen in the Hertzsprung-Russell diagram. In addition to colors and luminosity, these correlations
also involve metallicity in the case of main-sequence and red giant stars, and surface gravity in the case of white dwarfs.
For the youngest main-sequence stars, stellar age may play a role, too. Measured stellar colors are also affected
by interstellar dust extinction along the line of sight towards the star. Consequently, sufficiently accurate measurements of apparent
brightness and sufficient number of UV to IR colors, such as those that Rubin Observatory's Legacy Survey of Space
and Time (LSST; \citealt{2019ApJ...873..111I}) will provide, can be used to accurately estimate these parameters,
and ultimately stellar distances.
